---
title: "Simulation of limited commitment model"
author: "Mizuhiro Suzuki"
date: "6/4/2022"
bibliography: risksharing.bib
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages <- c(
  "tidyverse",
  "knitr",
  "nleqslv",
  "ggrepel",
  "rootSolve",
  "BB",
  "pracma",
  "latex2exp",
  "fixest"
)

pacman::p_load(packages, character.only = TRUE)

#knit: (function(input_file, encoding) {
#  out_dir <- 'docs';
#  rmarkdown::render(input = input_file,
#                   output_format = "html_document",
#                   encoding = encoding,
#                   output_dir = out_dir,
#                   output_file = 'index.html')})
#output: github_document
```


## Model settings

I simplify the model of @Laczo2015 and consider two households with $2 \times 2$ states.
The income process are iid:

$$
  y_1 = [2/3, 4/3],\ y_2 = [2/3, 4/3], \\
  \pi_1 = [0.1, 0.9],\ \pi_2 = [0.1, 0.9].
$$


```{r}
# income shocks and their transition probabilities of the household
inc1 <- c(2/3, 4/3)
P1 <- matrix(rep(c(0.1, 0.9), 2), nrow = 2, byrow = TRUE)

# income shocks and their transition probabilities of the village
inc2 <- c(2/3, 4/3)
P2 <- matrix(rep(c(0.1, 0.9), 2), nrow = 2, byrow = TRUE)

# transition super-matrix of income shocks
R <- kronecker(P2, P1)

# number of states
# (number of income states for a household times
# number of income states for the village)
S <- length(inc1) * length(inc2)

# Income matrix
# (col 1: HH income, col 2: village income)
inc_mat <- as.matrix(expand.grid(inc1, inc2))

# Aggregate income in each state
inc_ag <- rowSums(inc_mat)

```

I assume the CRRA utility functions:

$$
  u_i(c_{it}) = \frac{c_{it}^{1 - \sigma_i} - 1}{1 - \sigma_i}.
$$

```{r}
# Define utility function
util <- function(c, sigma) {
  if (sigma != 1) {
    output = (c ^ (1 - sigma) - 1) / (1 - sigma)
  } else if (sigma == 1) {
    output = log(c)
  }
  return(output)
}
util_prime <- function(c, sigma) c ^ (- sigma)
  
```

## Value of autarky

```{r}
delta <- 0.90 # time discount factor 
sigma1 <- 1.0 # coefficient of relative risk aversion of HH1
sigma2 <- 1.0 # coefficient of relative risk aversion of HH2
pcphi <- 0.0 # punishment under autarky

Uaut_func <- function(
  inc1, P1, inc2, P2,
  delta, sigma1, sigma2, pcphi,
  util, util_prime
  ) {

  # Expected utility under autarky for HH
  U1_aut <- numeric(length = length(inc1))
  i <- 1
  diff <- 1
  while (diff > 1e-12) {
    U1_aut_new <- util((inc1 * (1 - pcphi)), sigma1) + delta * P1 %*% U1_aut
    diff <- max(abs(U1_aut_new - U1_aut))
    U1_aut <- U1_aut_new
    i <- i + 1
  }

  # Expected utility under autarky for village
  U2_aut <- numeric(length = length(inc2))
  i <- 1
  diff <- 1
  while (diff > 1e-12) {
    U2_aut_new <- util((inc2 * (1 - pcphi)), sigma2) + delta * P2 %*% U2_aut
    diff <- max(abs(U2_aut_new - U2_aut))
    U2_aut <- U2_aut_new
    i <- i + 1
  }

  # Since in this case the income process are iid, 
  # values under autarky can be computed in the following way:
  # U1_out = util((1 - pcphi) * inc1, sigma1) + delta / (1 - delta) * P1 %*% util((1 - pcphi) * inc1, sigma1)
  # U2_out = util((1 - pcphi) * inc2, sigma2) + delta / (1 - delta) * P2 %*% util((1 - pcphi) * inc2, sigma2)
  # This way does not work when income processes are autocorrelated, as in the original model of Laczo (2015)

  # Matrix of expected utilities of autarky
  # (col 1: HH, col 2: village)
  Uaut <- expand.grid(U1_aut, U2_aut)

  return(Uaut)
}

Uaut <- Uaut_func(inc1, P1, inc2, P2, delta, sigma1, sigma2, pcphi, util, util_prime)

```

## Grid of relative Pareto weights and consumption on the grid points


```{r}
# (Number of grid points on relative Pareto weight) - 1
g <- 199

# The grid points of relative Pareto weights
qmin <- util_prime(max(inc2), sigma2) / util_prime(min(inc1 * (1 - pcphi)), sigma1)
qmax <- util_prime(min(inc2 * (1 - pcphi)), sigma2) / util_prime(max(inc1), sigma1)
q <- exp(seq(log(qmin), log(qmax), length.out = (g + 1)))

```

```{r}
# The grid points of consumption of HH 1
# Consumption is determined by aggregate income (inc_ag) and
# relative Pareto weights (q)
cons1 <- matrix(nrow = S, ncol = (g + 1))
for (k in 1:S) {
  for (l in 1:(g + 1)) {
    if (sigma1 == sigma2) {
      cons1[k, l] <- inc_ag[k] / (1 + q[l]^(- 1 / sigma1))
    } else {
      f = function(w) util_prime((inc_ag[k] - w), sigma2) - q[l] * util_prime(w, sigma1)
      v = uniroot(f, c(1e-5, (inc_ag[k] - 1e-5)), tol = 1e-8, maxiter = 100)
      cons1[k, l] <- v$root
    }
  }
}

```

## Values under risk-sharing (full)

```{r}

V_full_func <- function(
  inc1, P1, inc2, P2, g,
  R, S, inc_ag, qmin, qmax, q,
  delta, sigma1, sigma2, pcphi,
  util, util_prime, Uaut, cons1
  ) {

  # Value function iterations ================
  # Initial guess is expected utilities under autarky
  V1 <- outer(Uaut[, 1], rep(1, (g + 1)))
  V2 <- outer(Uaut[, 2], rep(1, (g + 1)))
  V1_new <- matrix(nrow = S, ncol = (g + 1))
  V2_new <- matrix(nrow = S, ncol = (g + 1))

  # Obtain value functions by value function iterations
  j <- 1
  diff <- 1
  while (diff > 1e-8 & j < 500) {
    V1_new <- util(cons1, sigma1) + delta * R %*% V1
    V2_new <- util(inc_ag - cons1, sigma2) + delta * R %*% V2
    diff <- max(max(abs(V1_new - V1)), max(abs(V2_new - V2)))
    V1 <- V1_new
    V2 <- V2_new
    j <- j + 1
  }

  V1_full <- V1
  V2_full <- V2

  return(list(V1_full, V2_full))
}

V_full <- V_full_func(
  inc1, P1, inc2, P2, g,
  R, S, inc_ag, qmin, qmax, q,
  delta, sigma1, sigma2, pcphi,
  util, util_prime, Uaut, cons1
)

V1_full <- V_full[[1]]
V2_full <- V_full[[2]]

```

## Values under risk-sharing (limited commitment)

```{r}

V_lc_func <- function(
  inc1, P1, inc2, P2, g,
  R, S, inc_ag, qmin, qmax, q,
  delta, sigma1, sigma2, pcphi,
  util, util_prime, Uaut, cons1, 
  V1_full, V2_full
  ) {

  # Value function iterations ================
  # Initial guess is expected utilities under full risk sharing
  V1 <- V1_full
  V2 <- V2_full
  V1_new <- matrix(nrow = S, ncol = (g + 1))
  V2_new <- matrix(nrow = S, ncol = (g + 1))

  diff <- 1
  iter <- 1
  maxiter <- 1000
  while ((diff > 1e-12) && (iter <= maxiter)) {

    cons1_lc <- cons1 # consumption of HH 1 under LC
    x_lc <- outer(rep(1, S), q)
    x_int <- matrix(nrow = S, ncol = 2) # matrix of bounds of intervals
  
    # First, ignore enforceability and just update the value functions
    # using the values at the previous iteration
    V1_new <- util(cons1, sigma1) + delta * R %*% V1
    V2_new <- util(inc_ag - cons1, sigma2) + delta * R %*% V2
    
    # Now check enforceability at each state
    for (k in 1:S) {
      
      # This function calculate the difference between the value when the relative Pareto weight is x
      # and the value of autarky. This comes from the equation (1) in the supplementary material of
      # Laczo (2015). Used to calculate the lower bound of the interval.
      calc_diff_val_aut_1 <- function(x) {
        # Calculate consumption of HH 1 at a relative Pareto weight x
        if (sigma1 == sigma2) {
          cons_x <- inc_ag[k] / (1 + x^(- 1 / sigma1))
        } else {
          f <- function(w) util_prime(inc_ag[k] - w, sigma2) / util_prime(w, sigma1) - x
          cons_x <- uniroot(f, c(1e-6, (inc_ag[k] - 1e-6)), tol = 1e-10, maxiter = 300)$root
        }
        
        # Index of the point on the grid of relative Pareto weights (q) closest to x
        q_ind <- which.min(abs(x - q))
        if (x > q[q_ind]) {
          q_ind <- q_ind + 1
        }
        
        # Value functions of HH 1 under the relative Pareto weights x
        # (I use interpolation since x might not be on the grid q)
        V1_x <- apply(V1, 1, function(y) approxfun(q, y, rule = 2)(x))
        
        # difference between 
        diff <- util(cons_x, sigma1) + delta * R[k,] %*% V1_x - Uaut[k, 1]
        return(diff)
      }
      
      # Similarly, this function is used to calculate the upper bound.
      calc_diff_val_aut_2 <- function(x) {
        # Calculate consumption of HH 1 at a relative Pareto weight x
        if (sigma1 == sigma2) {
          cons_x <- inc_ag[k] / (1 + x^(- 1 / sigma1))
        } else {
          f <- function(w) util_prime(inc_ag[k] - w, sigma2) / util_prime(w, sigma1) - x
          cons_x <- uniroot(f, c(1e-6, (inc_ag[k] - 1e-6)), tol = 1e-10, maxiter = 300)$root
        }
        
        # Index of the point on the grid of relative Pareto weights (q) closest to x
        q_ind <- which.min(abs(x - q))
        if (x > q[q_ind]) {
          q_ind <- q_ind + 1
        }
        
        # Value functions of HH 1 under the relative Pareto weights x
        # (I use interpolation since x might not be on the grid q)
        V2_x <- apply(V2, 1, function(y) approxfun(q, y, rule = 2)(x))
        
        # difference between 
        diff <- util(inc_ag[k] - cons_x, sigma2) + delta * R[k,] %*% V2_x - Uaut[k, 2]
        return(diff)
      }   
      
      # If the relative Pareto weight is too low and violates the PC, then
      # set the relative Pareto weight to the lower bound of the interval, and
      # HH1 gets the value under autarky.
      if ((calc_diff_val_aut_1(qmax) > 0) && calc_diff_val_aut_1(qmin) < 0) {
        
        # Derive the relative Pareto weight that satisfies the equation (1)
        # in the supplementary material of Laczo (2015)
        x_low <- uniroot(calc_diff_val_aut_1, c(qmin, qmax), tol = 1e-10, maxiter = 300)$root
        x_int[k, 1] <- x_low
        
        # Index of the point on the grid of relative Pareto weights (q) closest to x_low
        q_ind_low <- which.min(abs(x_low - q))
        if (x_low > q[q_ind_low]) {
          q_ind_low <- q_ind_low + 1
        }
        x_lc[k, 1:q_ind_low] <- x_low
        
        # Value functions of HH 2 under the relative Pareto weights x
        # (I use interpolation since x_low might not be on the grid q)
        V2_x <- apply(V2, 1, function(y) approxfun(q, y, rule = 2)(x_low))
        
        # Calculate consumption of HH 1 at a relative Pareto weight x_low
        if (sigma1 == sigma2) {
          cons_x_low <- inc_ag[k] / (1 + x_low^(- 1 / sigma1))
        } else {
          f <- function(w) util_prime(inc_ag[k] - w, sigma2) / util_prime(w, sigma1) - x_low
          cons_x_low <- uniroot(f, c(1e-6, (inc_ag[k] - 1e-6)), tol = 1e-10, maxiter = 300)$root
        }
        
        cons1_lc[k, 1:q_ind_low] <- cons_x_low
        V1_new[k, 1:q_ind_low] <- Uaut[k, 1]
        V2_new[k, 1:q_ind_low] <- util((inc_ag[k] - cons_x_low), sigma2) + delta * R[k,] %*% V2_x
      } else if (calc_diff_val_aut_1(qmax) <= 0) {
        V1_new[k,] <- Uaut[k, 1]
        V2_new[k,] <- Uaut[k, 2]
        x_int[k, 1] <- qmax
        x_lc[k,] <- qmax
        q_ind_low <- g + 2
      } else if (calc_diff_val_aut_1(qmin) >= 0) {
        x_int[k, 1] <- qmin
        q_ind_low <- 0
      }
    
      # If the relative Pareto weight is too high and violates the PC, then
      # set the relative Pareto weight to the upper bound of the interval, and
      # HH2 gets the value under autarky.
      if (q_ind_low <= (g + 1)) {
        if ((calc_diff_val_aut_2(qmin) > 0) && calc_diff_val_aut_2(qmax) < 0) {
          
          # Derive the relative Pareto weight that satisfies the equation (2)
          # in the supplementary material of Laczo (2015)
          x_high <- uniroot(calc_diff_val_aut_2, c(qmin, qmax), tol = 1e-10, maxiter = 300)$root
          x_int[k, 2] <- x_high
          
          # Index of the point on the grid of relative Pareto weights (q) closest to x_low
          q_ind_high <- which.min(abs(x_high - q))
          if (x_high > q[q_ind_high]) {
            q_ind_high <- q_ind_high + 1
          }
          x_lc[k, q_ind_high:(g + 1)] <- x_high
          
          # Value functions of HH 2 under the relative Pareto weights x
          # (I use interpolation since x_high might not be on the grid q)
          V1_x <- apply(V1, 1, function(y) approxfun(q, y, rule = 2)(x_high))
          
          # Calculate consumption of HH 1 at a relative Pareto weight x_high
          if (sigma1 == sigma2) {
            cons_x_high <- inc_ag[k] / (1 + x_high^(- 1 / sigma1))
          } else {
            f <- function(w) util_prime(inc_ag[k] - w, sigma2) / util_prime(w, sigma1) - x_high
            cons_x_high <- uniroot(f, c(1e-6, (inc_ag[k] - 1e-6)), tol = 1e-10, maxiter = 300)$root
          }
          
          cons1_lc[k, q_ind_high:(g + 1)] <- cons_x_high
          V1_new[k, q_ind_high:(g + 1)] <- util(cons_x_high, sigma1) + delta * R[k,] %*% V1_x
          V2_new[k, q_ind_high:(g + 1)] <- Uaut[k, 2]
        } else if (calc_diff_val_aut_2(qmin) <= 0) {
          V1_new[k,] <- Uaut[k, 1]
          V2_new[k,] <- Uaut[k, 2]
          x_int[k, 2] <- qmin
          x_lc[k,] <- qmin
          q_ind_high <- 0
        } else if (calc_diff_val_aut_2(qmax) >= 0) {
          x_int[k, 2] <- qmax
          q_ind_high <- g + 2
        }
      }
      
      # The case where the relative Pareto weight does not violate PC
      if ((q_ind_low + 1) < (q_ind_high - 1)) {
        V1_new[k, (q_ind_low + 1):(q_ind_high - 1)] <- 
          util(cons1[k, (q_ind_low + 1):(q_ind_high - 1)], sigma1) + 
          delta * R[k,] %*% V1[, (q_ind_low + 1):(q_ind_high - 1)]
        V2_new[k, (q_ind_low + 1):(q_ind_high - 1)] <- 
          util(inc_ag[k] - cons1[k, (q_ind_low + 1):(q_ind_high - 1)], sigma2) + 
          delta * R[k,] %*% V2[, (q_ind_low + 1):(q_ind_high - 1)]
      }
      
    }
    
    diff <- max(max(abs(V1_new - V1)), max(abs(V2_new - V2)))
    V1 <- V1_new
    V2 <- V2_new
    iter <- iter + 1
    
  }

  if (iter == maxiter) {
    print("Reached the maximum limit of iterations!")
  }

  return(list(V1, V2, x_int))
}

```

```{r, include = FALSE, eval = FALSE}

V_lc <- V_lc_func(
  inc1, P1, inc2, P2, g,
  R, S, inc_ag, qmin, qmax, q,
  delta, sigma1, sigma2, pcphi,
  util, util_prime, Uaut, cons1, 
  V1_full, V2_full
  )

V1_lc <- V_lc[[1]]
V2_lc <- V_lc[[2]]
x_int_lc <- V_lc[[3]]
log(x_int_lc)

```

## Make a pipeline based on the functions defined above

```{r}

V_lc_all_func <- function(
  inc1, P1, inc2, P2, g,
  delta, sigma1, sigma2, pcphi,
  util, util_prime
  ) {

  # Pre settings ====================
  # transition super-matrix of income shocks
  R <- kronecker(P2, P1)

  # number of states
  # (number of income states for a household times
  # number of income states for the village)
  S <- length(inc1) * length(inc2)

  # Aggregate income in each state
  inc_ag <- rowSums(expand.grid(inc1, inc2))

  # The grid points of relative Pareto weights
  qmin <- util_prime(max(inc2), sigma2) / util_prime(min(inc1 * (1 - pcphi)), sigma1)
  qmax <- util_prime(min(inc2 * (1 - pcphi)), sigma2) / util_prime(max(inc1), sigma1)
  q <- exp(seq(log(qmin), log(qmax), length.out = (g + 1)))

  # The grid points of consumption of HH 1
  # Consumption is determined by aggregate income (inc_ag) and
  # relative Pareto weights (q)
  cons1 <- matrix(nrow = S, ncol = (g + 1))
  for (k in 1:S) {
    for (l in 1:(g + 1)) {
      if (sigma1 == sigma2) {
        cons1[k, l] <- inc_ag[k] / (1 + q[l]^(- 1 / sigma1))
      } else {
        f = function(w) util_prime((inc_ag[k] - w), sigma2) - q[l] * util_prime(w, sigma1)
        v = uniroot(f, c(1e-5, (inc_ag[k] - 1e-5)), tol = 1e-8, maxiter = 100)
        cons1[k, l] <- v$root
      }
    }
  }

  # Matrix of expected utilities of autarky
  # (col 1: HH, col 2: village)
  Uaut <- Uaut_func(inc1, P1, inc2, P2, delta, sigma1, sigma2, pcphi, util, util_prime)

  # Values under full risk-sharing
  V_full <- V_full_func(
    inc1, P1, inc2, P2, g,
    R, S, inc_ag, qmin, qmax, q,
    delta, sigma1, sigma2, pcphi,
    util, util_prime, Uaut, cons1
  )

  V1_full <- V_full[[1]]
  V2_full <- V_full[[2]]

  # Values under risk-sharing with dynamic limited commitment
  V_lc <- V_lc_func(
    inc1, P1, inc2, P2, g,
    R, S, inc_ag, qmin, qmax, q,
    delta, sigma1, sigma2, pcphi,
    util, util_prime, Uaut, cons1,
    V1_full, V2_full
    )

  V1_lc <- V_lc[[1]]
  V2_lc <- V_lc[[2]]
  x_int_lc <- V_lc[[3]]

  return(list(V1_full, V2_full, V1_lc, V2_lc, x_int_lc))
}

```

## Simulation


First solve the model:

```{r}
output <- V_lc_all_func(
   inc1, P1, inc2, P2, g, 
   delta, sigma1, sigma2, pcphi, util, util_prime
)

```

Start from the "fair" Pareto weight:

```{r}

q_start = q[which.min(abs(q - 1.0))]

```

```{r}

N_sim <- 500

income_realization <- rep(0, N_sim)

income_realization[1] <- sample(seq(S), size = 1, prob = R[1,])
for (i in 2:N_sim) {
  income_realization[i] <- sample(seq(S), size = 1, prob = R[income_realization[i - 1],])
}

income_realization <- sample(
  seq(S),
  size = N_sim,
  replace = TRUE,
  prob = R[1,]
)

income_realization[1]

q_int <- output[[5]]

q_sim <- rep(0, N_sim + 1)
q_sim[1] <- q_start
for (t in seq(N_sim)) {
  q_int_temp <- q_int[income_realization[t],]
  if (q_sim[t] < q_int_temp[1]) {
    q_sim[t + 1] <- q_int_temp[1]
  } else if (q_sim[t] > q_int_temp[2]) {
    q_sim[t + 1] <- q_int_temp[2]
  } else {
    q_sim[t + 1] <- q_sim[t]
  }
}

consumption_realization_lc <- matrix(NA, nrow = N_sim + 1, ncol = 2)
for (t in 2:(N_sim + 1)) {
  f <- function(w) util_prime(inc_ag[income_realization[t - 1]] - w, sigma2) / util_prime(w, sigma1) - q_sim[t]
  consumption_realization_lc[t, 1] <- uniroot(f, c(1e-6, (inc_ag[income_realization[t - 1]] - 1e-6)), tol = 1e-10, maxiter = 300)$r
  consumption_realization_lc[t, 2] <- inc_ag[income_realization[t - 1]] - consumption_realization_lc[t, 1]
}

```

```{r}

df_lc <- bind_rows(
  tibble(
    income = inc_mat[income_realization, 1],
    consumption = consumption_realization_lc[2:(N_sim + 1), 1],
    time = seq(N_sim),
    id = 1
  ),
  tibble(
    income = inc_mat[income_realization, 2],
    consumption = consumption_realization_lc[2:(N_sim + 1), 2],
    time = seq(N_sim),
    id = 2
  )
)

feols(consumption ~ income | id + time, df_lc)






```


